A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2022-05-21 review/response.Rmd

Thank you for the review of our article. These are the changes that we have made in response. The review comment is indented italic, and our response follows in regular text.

First, we have reformatted the paper to use the `rjtools` package.

We have made a simpler code example that will run from start to end of the algorithm, but also kept the full Australia example created from data provided with the source files of the article.

> *The article is clearly written and does a good job of discussing previous mapping approaches as well as their limitations. The figures were useful in this regard, as they make the textual discussion in the background sections easily understandable even for users less experienced in geography. The flowchart summarizing the algorithm was also useful. I would suggest some minor substantive revisions to improve the substance of the background as well as the clarity of the package’s functionality. First, the discussion of the pros and cons of choropleth maps and various forms of cartograms was good, but I would like more information about the value of hexagons over other shapes. Why not circles or squares or triangles? Hexagons do have advantages, but adding a few more sentences about this would be useful, especially for non-specialist users. Second, other R packages for mapping can be cited, including cartogram (Jeworutzki et al 2020) and tilemaps (McNeill and Hale 2017). *

Thank you for the affirmation of the use of written and visual descriptions. A small paragraph has been included at the end of Existing Mapping Practices to clearly describe the benefits of the connectivity of hexagons as per your suggestion.
References to the R package implementations of other Mapping Practice have now been made explicit.

> *Third, I might recommend putting the “implementation” section above the “user choices” section. Fourth, and this is probably a personal preference, the implementation section could be improved by adding a simple table listing the main functions in the package and a brief summary of what they do. This makes the information more easily accessible to the reader. Fifth, the implementation section provides a running example, but please include some sample code directly in the article. This makes the package more accessible to users. It is not necessary to include all code of course, but some of the more important functions that produce a core visualization should be provided directly in the text.*

@@

> *There were also a few minor typos that should be fixed (e.g., “tyroid” in the caption of Figure 3 and the comma beginning the third line of the second paragraph of the “implementation” section). The last two sentences of the first paragraph of the discussion section are also identical to the next two sentences. Overall, these revisions to the article are relatively minor and manageable.*

> *The package itself seems to be reasonably well-organized, the code is clean, and there is sufficient built-in documentation. The progress bars are also useful. I don’t have any major suggestions for improving the functionality of the package itself, other than the fact that I couldn’t reproduce the example maps in the implementation section because the data were not self-contained. The article says that the example is reproducible using data from the absmapsdata package, but in the provided RMD file the object “fort_aus” can be traced back to the “SIR Downloadable Data.csv” dataset, which is from an external source (and also not available on the github repo). Forgive me if I am missing something here (and it is possible that I am). If not, these SIR data should be embedded or example code fixed to not be dependent on the SIR data so that the illustrative maps are reproducible.*

The implementation section now focusses only on Tasmania, for the sake of speed.
We have now included the R data file for a completed Australian thyroid hexagon map.
There are two sets of data required to make the Australian thyroid map, the shape file and SIR data.
The SIR Downloadable Data csv file can be found in the data folder.

### 2022-06-02 review/response.Rmd

> *There were also a few minor typos that should be fixed (e.g., “tyroid” in the caption of Figure 3 and the comma beginning the third line of the second paragraph of the “implementation” section). The last two sentences of the first paragraph of the discussion section are also identical to the next two sentences. Overall, these revisions to the article are relatively minor and manageable.*

Fixed, and paper has been carefully proofread again.

> *The package itself seems to be reasonably well-organized, the code is clean, and there is sufficient built-in documentation. The progress bars are also useful. I don’t have any major suggestions for improving the functionality of the package itself, other than the fact that I couldn’t reproduce the example maps in the implementation section because the data were not self-contained. The article says that the example is reproducible using data from the absmapsdata package, but in the provided RMD file the object “fort_aus” can be traced back to the “SIR Downloadable Data.csv” dataset, which is from an external source (and also not available on the github repo). Forgive me if I am missing something here (and it is possible that I am). If not, these SIR data should be embedded or example code fixed to not be dependent on the SIR data so that the illustrative maps are reproducible.*

For improving the speed to knit the paper the implementation section uses only the small data set corresponding to Tasmania.

The data files for the full Australia thyroid choropleth and hexagon map are provided so that the Figures 1 and 2 can be produced quickly. The code to produce these data sets is provided, so that the interested reader can follow the entire hexagon map generating process. There are two sets of data required to make the Australian thyroid map, the shape file and thyroid SIR data. The `SIR Downloadable Data.csv` file is located in the data folder. The shape file is the `sa22011` data from the `absmapsdata` package, which would need to be installed with `remotes::install_github("wfmackey/absmapsdata")`.

### 2022-06-02 kobakian-cook.Rmd
links now working and more tweakig of paper

Over time, the density of living in major cities has increased as residents have gathered to live near urban areas [@ACTUC]. The division of these residents into approximately equal population areas results in dramatically different square meterage of administrative geographic units, such as states or electorates. In a choropleth map display, the geographic units are coloured to represent the value of the statistic for each unit [@EI]. This can cause some values to be emphasised over others, and allows choropleth map displays to misrepresent the spatial distributions of human related statistics due to area-size bias [@BCM]. Figure \ref{fig:choro} is a choropleth map that uses colour to display the estimated standardised incidence ratios (SIRs) of thyroid cancer for females, in each of the Statistical Areas at Level 2 (SA2) used by the @abs2011. The Australian choropleth map display draws attention to the expanse of dark and light blue areas across the rural communities in all states. The SA2s on the east coast around Brisbane and in northern New South Wales stand out as more orange and red. However, this display neglects the vast amount of Australia residents living in the densely populated capital cities.

@@

This paper contains a discussion of Existing Mapping Practices. This is followed by details of the Algorithm. The implementation section discusses the algorithm in the \CRANpkg{sugarbag} package. Finally, the use and benefits of animation are presented.

## Existing Mapping Practices

@@

Administrative areas are often used to aggregate census data and used to understand demographics within communities of the Australian population. Each SA2 [@abs2011] was designed to represent a community. This collection of communities presents an opportunity for explicit map transformations to improve communication of spatial distributions [@CBATCC]. In Figure \ref{fig:tas_displays} the choropleth map can be seen underneath each alternative map display to allow for comparisons to be made.

There are several established alternative visualisation methods. Rectangular cartograms [@ORC] and Dorling cartograms [@ACTUC] implemented in \pkg{cartogram} and tile maps implemented in \CRANpkg{tilemaps}, all use one simple shape to represent each geographic unit. They all minimise the emphasis on the size or shape of the geographic areas. These alternative map displays focus on the relationship between neighbours, attempting to preserve connections, and disregard the unique shapes of the administrative boundaries. Figure \ref{fig:tas_displays} shows a collection of alternative map displays, this includes a) a contiguous cartogram, b) a non-contiguous cartogram, and c) a Dorling cartogram.

@@

tas_ggcont <- ggplot(cont) + 
  geom_sf(fill = "#D3D3D3", colour = "white", data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = "grey", size = 0.01) + 
  scale_fill_identity() + guides(fill = "none") +
  theme_map() + ggtitle("a")
#tas_ggcont
```

@@

tas_ggncont <- ggplot(ncont %>% filter(state_name_2011 == "Tasmania")) + 
  geom_sf(fill = "#D3D3D3", colour = "white",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  guides(fill = "none") +
  theme_map() + ggtitle("b")
#tas_ggncont
```

```{r dorl}
# Non - Contiguous Dorling Cartograms
dorl <- sa2thyroid_ERP %>% 
  filter(!is.na(Population)) %>% 
  #mutate(pop = (Population/max(Population))*10) %>% # helpful for all of Australia
  cartogram_dorling(., weight = "Population", k = 0.2, m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
tas_ggdorl <- ggplot(dorl) +
  geom_sf(fill = "#D3D3D3", colour = "white",  data = sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR), colour = "grey", size = 0.01) + 
  scale_fill_identity() +
  guides(fill = "none") +
  theme_map() + ggtitle("c")
#tas_ggdorl
```

```{r tas_displays, width = 10, height = 3, fig.cap = "The three displays show alternative maps of the Australian state of Tasmania at SA2 level: (a) contiguous cartogram, (b) non-contiguous cartogram and (c) Dorling cartogram of Tasmania. The contiguous cartogram looks like the state has an hourglass figure, while the non-contiguous cartogram shrinks areas into invisibility. The Dorling expands the metropolitan regions."}
library(gridExtra)
grid.arrange(tas_ggcont, tas_ggncont, tas_ggdorl, ncol=3)
```

When communicating information that is relevant to the population, each member of the population can be given equal representation by transforming the map [@TVSSS]. The connected communities can be recognised by allowing the boundaries to maintain connection in the transformed display. The contiguous cartogram displayed in Figure \ref{fig:tas_displays}a) draws attention to smaller geographic units when they are rescaled according to the population [@DMAHP]. These new shapes can now be coloured to represent a second variable. The algorithm uses the geographic shape of the areas and iterates toward sizing the areas to represent the population. This display can create twisted and unfamiliar shapes from the geographic units as the algorithms must satisfy the topology conditions, especially when there are communities located geographically far from their neighbours [@TVSSS].

@@

The non-contiguous cartogram in Figure \ref{fig:tas_displays}b) also uses the population to rescale the geographic units. Unlike the contiguous cartogram, the SA2 areas maintain their geographic shape, but they may not retain the connection with their neighbours. The population of the SA2 areas is used to scale the geographic units in the non-contiguous cartogram [@NAC]. The amount of background space can be meaningful in non-contiguous cartograms [@ECGC]. The size disparity between rural areas and urban areas result in reasonable display within the south eastern city of Hobart, but these units are not visible in the context of the Tasmania state map. Depending on the difference between the population and geographic land size, the amount of white space can also prevent meaningful understanding of the distribution [@TVSSS].

@@

The Dorling cartogram presents each geographic unit as a circle, the size of the circle is scaled according to the population value of each area [@ACTUC]. Figure \ref{fig:tas_displays}c) shows the Tasmania SA2 areas as an individual circle located as close as possible to the geographic centroid location. This map draws attention to the collection of coastal cities in Tasmania that were not apparent in Figure \ref{fig:tas_displays} a), or b). This display also highlights that the approximation of equal population is not easy to fulfil, as there is some disparity in the sizes of the circles.

@@

The result of the algorithm applied to the same data represented in Figure \ref{fig:choro} shows rates of thyroid cancer for females in each of the SA2 areas of Australia as hexagons in Figure \ref{fig:hexmap}.
The difference to the choropleth map in Figure \ref{fig:choro} are clear. Now the higher thyroid cancer incidence in the densely populated areas in Sydney, Brisbane and Perth are visible. Interestingly, there is no clear divide between rural and urban SA2 areas, as many rural areas and the cities of Melbourne, Darwin, Adelaide and Hobart have low rates of thyroid incidence for females. The hexagon tile map display provides a more accurate representation of the spatial distribution of thyroid cancer incidence across Australia.

Figure \ref{fig:hexmap} highlights the density of Australian capital cities, as it draws attention to the many communities in Sydney, Melbourne and Hobart. This display also highlights the disparity in the burden of thyroid cancer for females in the communities of these cities. There are several collections of red hexagons in Sydney that represent the communities with much higher rates of diagnosis than the Australian average. Brisbane also experiences higher than average rates of diagnosis, but has more orange than red. The females in the cities of Adelaide and Perth  show much lower rates of diagnosis.

Compared to the choropleth map display in Figure \ref{fig:choro}, the low rates in the rural Australian communities are no longer dominating the display. While the representative hexagons are still visible against the black background, the much lower than average rates in rural Australia are less noticeable.

@@

There are several key steps in the creation of the hexagon tile map as described in the flow chart in Figure \ref{fig:sugarbag_flow}. First, derive the set of centroids from the polygons provided, then create the grid of hexagon locations. These two processes are defined in the blue left column of the flow chart in figure \ref{fig:sugarbag_flow}. Each centroid can then be allocated to an available hexagon location. The steps for the allocation process are detailed in the right column of Figure \ref{fig:sugarbag_flow}. There are several filter steps to speed the process of selecting an appropriate hexagon to represent each geographic unit. To make tessellated plots with the hexagon allocations, the point locations are converted into hexagon shapes.

\begin{figure}
\includegraphics[width=14cm]{figs/sugarbag flow.png}
\caption{\label{fig:sugarbag_flow}A flow diagram detailing the steps taken to create a hexagon tile map. There are two basic processes, one to make the grid, and the other to allocate centroids to grid points.}
\end{figure}

@@

The results from various steps in the process are illustrated in Figure \ref{fig:filterprocess}.

@@

A grid is created to allow tessellation of the hexagons that represent the geographic units. For a hexagon tile map, the grid of possible hexagon locations is made using the \code{create\_grid} function. It uses the centroids, the hexagon size and the buffer distance. Figure \ref{fig:filterprocess} 1a) shows the initial locations of grid points created for Tasmania.

@@

A set of longitude columns, and latitude rows are created to define the locations of the hexagons. The distance between each row and column is the size specified by \code{hex\_size}. Equally spaced columns are created from the minimum longitude minus the buffer distance, up to the maximum longitude plus the buffer distance. Similarly, the rows are created from the latitude values and the buffer distance. A unique hexagon location is created from all intersections of the longitude columns and latitude rows. Figure \ref{fig:filterprocess} 1a) shows the hexagon grid after every second latitude row on the grid is shifted right, by half of the hexagon size.

@@

The hexagon grid points are kept only if they fall within the rolling average of the minimum and maximum centroid values after accounting for the buffer distance, for each row and column of the grid. Figure \ref{fig:filterprocess} 1b) displays remaining hexagon grid points after applying the buffer filter. The sparsely populated South-West region of National Park has fewer points available in the water compared to the South-East region near the city of Hobart.

@@

The distance is calculated between each centroid in the set, and each of the focal points provided. The order for allocation is determined by the distance between the polygon centroid and it's closest focal point. For this example, this distance is only calculated for the capital city of Hobart, represented in figure \ref{fig:filterprocess} 2a) and 2b) as a white cross.

@@

The set of possible hexagon grid points reduces by one after each centroid is allocated, then only the grid points that have not yet been selected are considered. Keeping only the available hexagon points prevents multiple geographic units from being allocated to the same hexagon. This is demonstrated in figure \ref{fig:filterprocess} 2a) and 2b) by the black hexagons that represent the seven closest polygons to the capital city of Hobart. As the allocation process begins for the eighth closest centroid there are seven unavailable hexagon locations.

@@

The \code{width} parameter is used to take a slice of the remaining points. The width is the amount of degrees used on either side of the angle from the focal point to centroid location. This uses the angle from the closest capital city, to the current centroid as seen in figure \ref{fig:filterprocess} 2a). This allows the spatial relationship to be preserved, even when it is allocated to a hexagon that is further from the focal point then the original centroid location.

@@

This algorithm should be general enough to apply to other study regions and contexts. The algorithm has only been tested using single countries. While the buffer allows extension beyond the furthest centroids, there is no mechanism implemented to protect the borders and ensure centroids are placed within the geographic borders of a country.
The algorithm is also currently limited to representing each geographic area with only one hexagon. Future developments may address these limitations, and refine the filter steps for final selection of a hexagon grid point for each centroid. A direct angle is currently implemented, but may narrowly miss appropriate hexagons, a logarithmic function may help to choose a closer hexagon to the original centroid location.
This algorithm is an effective start to automatically creating hexagon tile maps for many geographic units. This contributes to an extensive body of work that encourages the use of alternative map displays for statistical displays of spatial data.

>>>>

Over time, the density of living in major cities has increased as residents have gathered to live near urban areas [@ACTUC]. The division of these residents into approximately equal population areas results in dramatically different square meterage of administrative geographic units, such as states or electorates. In a choropleth map display, the geographic units are coloured to represent the value of the statistic for each unit [@EI]. This can cause some values to be emphasised over others, and allows choropleth map displays to misrepresent the spatial distributions of human related statistics due to area-size bias [@BCM]. Figure \@ref(fig:choro) is a choropleth map that uses colour to display the estimated standardised incidence ratios (SIRs) of thyroid cancer for females, in each of the Statistical Areas at Level 2 (SA2) used by the @abs2011. The Australian choropleth map display draws attention to the expanse of dark and light blue areas across the rural communities in all states. The SA2s on the east coast around Brisbane and in northern New South Wales stand out as more orange and red. However, this display neglects the vast amount of Australia residents living in the densely populated capital cities.

@@

This paper contains a discussion of existing mapping practices, followed by details of the algorithm. The implementation section explains the algorithm available in the \CRANpkg{sugarbag} package. How to utilise animation with the hexagon map is described. The paper finishes with a summary and possible new directions for the algorithm.

## Existing mapping practices

@@

Administrative areas are often used to aggregate census data and used to understand demographics within communities of the Australian population. Each SA2 [@abs2011] was designed to represent a community. This collection of communities presents an opportunity for explicit map transformations to improve communication of spatial distributions [@CBATCC]. In Figure \@ref(fig:tasdisplays) the choropleth map can be seen underneath each alternative map display to allow for comparisons to be made.

There are several established alternative visualisation methods. Rectangular cartograms [@ORC] and Dorling cartograms [@ACTUC] implemented in \pkg{cartogram} and tile maps implemented in \CRANpkg{tilemaps}, all use one simple shape to represent each geographic unit. They all minimise the emphasis on the size or shape of the geographic areas. These alternative map displays focus on the relationship between neighbours, attempting to preserve connections, and disregard the unique shapes of the administrative boundaries. Figure \@ref(fig:tasdisplays) shows a collection of alternative map displays, this includes a) a contiguous cartogram, b) a non-contiguous cartogram, and c) a Dorling cartogram.

@@

tas_ggcont <- ggplot(cont) + 
  geom_sf(fill = "#D3D3D3", colour = "white", data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = "grey", size = 0.01) + 
  scale_fill_identity() + guides(fill = "none") +
  theme_map() #+ ggtitle("a")
#tas_ggcont
```

@@

tas_ggncont <- ggplot(ncont %>% filter(state_name_2011 == "Tasmania")) + 
  geom_sf(fill = "#D3D3D3", colour = "white",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  guides(fill = "none") +
  theme_map() #+ ggtitle("b")
#tas_ggncont
```

```{r dorl}
# Non - Contiguous Dorling Cartograms
dorl <- sa2thyroid_ERP %>% 
  filter(!is.na(Population)) %>% 
  #mutate(pop = (Population/max(Population))*10) %>% # helpful for all of Australia
  cartogram_dorling(., weight = "Population", k = 0.2, m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
tas_ggdorl <- ggplot(dorl) +
  geom_sf(fill = "#D3D3D3", colour = "white",  data = sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR), colour = "grey", size = 0.01) + 
  scale_fill_identity() +
  guides(fill = "none") +
  theme_map() #+ ggtitle("c")
#tas_ggdorl
```

```{r tasdisplays, out.width="100%", fig.width = 8, fig.height = 3, layout = "l-body", fig.cap = "The three displays show alternative maps of the Australian state of Tasmania at SA2 level: (a) contiguous cartogram, (b) non-contiguous cartogram and (c) Dorling cartogram of Tasmania. The contiguous cartogram looks like the state has an hourglass figure, while the non-contiguous cartogram shrinks areas into invisibility. The Dorling expands the metropolitan regions."}
#library(gridExtra)
#grid.arrange(tas_ggcont, tas_ggncont, tas_ggdorl, ncol=3)
cowplot::plot_grid(tas_ggcont, 
                   tas_ggncont, 
                   tas_ggdorl, ncol=3,
                   labels = c("a", "b", "c"))

When communicating information that is relevant to the population, each member of the population can be given equal representation by transforming the map [@TVSSS]. The connected communities can be recognised by allowing the boundaries to maintain connection in the transformed display. The contiguous cartogram displayed in Figure \@ref(fig:tasdisplays)a draws attention to smaller geographic units when they are rescaled according to the population [@DMAHP]. These new shapes can now be coloured to represent a second variable. The algorithm uses the geographic shape of the areas and iterates toward sizing the areas to represent the population. This display can create twisted and unfamiliar shapes from the geographic units as the algorithms must satisfy the topology conditions, especially when there are communities located geographically far from their neighbours [@TVSSS].

@@

The non-contiguous cartogram in Figure \@ref(fig:tasdisplays)b also uses the population to rescale the geographic units. Unlike the contiguous cartogram, the SA2 areas maintain their geographic shape, but they may not retain the connection with their neighbours. The population of the SA2 areas is used to scale the geographic units in the non-contiguous cartogram [@NAC]. The amount of background space can be meaningful in non-contiguous cartograms [@ECGC]. The size disparity between rural areas and urban areas result in reasonable display within the south eastern city of Hobart, but these units are not visible in the context of the Tasmania state map. Depending on the difference between the population and geographic land size, the amount of white space can also prevent meaningful understanding of the distribution [@TVSSS].

@@

The Dorling cartogram presents each geographic unit as a circle, the size of the circle is scaled according to the population value of each area [@ACTUC]. Figure \@ref(fig:tasdisplays)c shows the Tasmania SA2 areas as an individual circle located as close as possible to the geographic centroid location. This map draws attention to the collection of coastal cities in Tasmania that were not apparent in Figure \@ref(fig:tasdisplays) a or b. This display also highlights that the approximation of equal population is not easy to fulfil, as there is some disparity in the sizes of the circles.

@@

The result of the algorithm applied to the same data represented in Figure \@ref(fig:choro) shows rates of thyroid cancer for females in each of the SA2 areas of Australia as hexagons in Figure \@ref(fig:hexmap).
The difference to the choropleth map in Figure \@ref(fig:choro) are clear. Now the higher thyroid cancer incidence in the densely populated areas in Sydney, Brisbane and Perth are visible. Interestingly, there is no clear divide between rural and urban SA2 areas, as many rural areas and the cities of Melbourne, Darwin, Adelaide and Hobart have low rates of thyroid incidence for females. The hexagon tile map display provides a more accurate representation of the spatial distribution of thyroid cancer incidence across Australia.

Figure \@ref(fig:hexmap) highlights the density of Australian capital cities, as it draws attention to the many communities in Sydney, Melbourne and Hobart. This display also highlights the disparity in the burden of thyroid cancer for females in the communities of these cities. There are several collections of red hexagons in Sydney that represent the communities with much higher rates of diagnosis than the Australian average. Brisbane also experiences higher than average rates of diagnosis, but has more orange than red. The females in the cities of Adelaide and Perth  show much lower rates of diagnosis.

Compared to the choropleth map display in Figure \@ref(fig:choro), the low rates in the rural Australian communities are no longer dominating the display. While the representative hexagons are still visible against the black background, the much lower than average rates in rural Australia are less noticeable.

@@

There are several key steps in the creation of the hexagon tile map as described in the flow chart in Figure \@ref(fig:sugarbagflow). First, derive the set of centroids from the polygons provided, then create the grid of hexagon locations. These two processes are defined in the blue left column of the flow chart in figure \@ref(fig:sugarbagflow). Each centroid can then be allocated to an available hexagon location. The steps for the allocation process are detailed in the right column of Figure \@ref(fig:sugarbagflow). There are several filter steps to speed the process of selecting an appropriate hexagon to represent each geographic unit. To make tessellated plots with the hexagon allocations, the point locations are converted into hexagon shapes.

```{r sugarbagflow, out.width = "100%", layout = "l-body-outset", fig.cap = "A flow diagram detailing the steps taken to create a hexagon tile map. There are two basic processes, one to make the grid, and the other to allocate centroids to grid points.", fig.alt="The steps of the algorithm to creating a hexagon grid with buffer, and the allocation of spatial polygon centroids to hexagon. This is the same information as provided in the text of the paper."}
knitr::include_graphics("figs/sugarbag_flow.png")
```

@@

The results from various steps in the process are illustrated in Figure \@ref(fig:filterprocess).

@@

A grid is created to allow tessellation of the hexagons that represent the geographic units. For a hexagon tile map, the grid of possible hexagon locations is made using the \code{create\_grid} function. It uses the centroids, the hexagon size and the buffer distance. Figure \@ref(fig:filterprocess) 1a shows the initial locations of grid points created for Tasmania.

@@

A set of longitude columns, and latitude rows are created to define the locations of the hexagons. The distance between each row and column is the size specified by \code{hex\_size}. Equally spaced columns are created from the minimum longitude minus the buffer distance, up to the maximum longitude plus the buffer distance. Similarly, the rows are created from the latitude values and the buffer distance. A unique hexagon location is created from all intersections of the longitude columns and latitude rows. Figure \@ref(fig:filterprocess) 1a shows the hexagon grid after every second latitude row on the grid is shifted right, by half of the hexagon size.

@@

The hexagon grid points are kept only if they fall within the rolling average of the minimum and maximum centroid values after accounting for the buffer distance, for each row and column of the grid. Figure \@ref(fig:filterprocess) 1b displays remaining hexagon grid points after applying the buffer filter. The sparsely populated South-West region of National Park has fewer points available in the water compared to the South-East region near the city of Hobart.

@@

The distance is calculated between each centroid in the set, and each of the focal points provided. The order for allocation is determined by the distance between the polygon centroid and it's closest focal point. For this example, this distance is only calculated for the capital city of Hobart, represented in Figure \@ref(fig:filterprocess) 2a and 2b as a white cross.

@@

The set of possible hexagon grid points reduces by one after each centroid is allocated, then only the grid points that have not yet been selected are considered. Keeping only the available hexagon points prevents multiple geographic units from being allocated to the same hexagon. This is demonstrated in Figure \@ref(fig:filterprocess) 2a and 2b by the black hexagons that represent the seven closest polygons to the capital city of Hobart. As the allocation process begins for the eighth closest centroid there are seven unavailable hexagon locations.

@@

The \code{width} parameter is used to take a slice of the remaining points. The width is the amount of degrees used on either side of the angle from the focal point to centroid location. This uses the angle from the closest capital city, to the current centroid as seen in Figure \@ref(fig:filterprocess) 2a. This allows the spatial relationship to be preserved, even when it is allocated to a hexagon that is further from the focal point then the original centroid location.

@@

Australia is used as an example, because it represents an extreme sparse geography and small dense population areas which produces unsatisfactory results with all existing cartogram-type algorithms. However, the algorithm is sufficiently general to apply to other geographic regions, and indeed has been applied to make hexagon maps of the UK, USA and France.

There are many possible extensions to the algorithm that might be valuable future work. The buffer allows extension beyond the furthest centroids, but there is no mechanism to protect borders to force centroids are placed within the geographic borders of a region (e.g. country or state). The algorithm is also currently limited to representing each geographic area with only one hexagon. In the filter step for final selection of a hexagon grid point for each centroid, a direct angle constraint is currently implemented, but it may narrowly miss potentially better hexagon allocation. Some other constraints, such as a logarithmic function may help to choose refine the final allocation to hexagons closer to the original centroid location. In a country like Australia, the vast centre produces a hexagon map with small isolated hexagons, even when the city centres are hugely expanded. It could be useful to allow variable sized hexagons, primarily to increase the size of the isolated hexagons to improve their visibility.

This new algorithm provides an effective start point for automatically creating hexagon tile maps for any spatial polygon data, and contributes to an extensive body of work that encourages the use of alternative map displays for statistical displays of spatial data.
